Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-07 volunteer-clouds-ieee-ci-si.tex
Adding correlation

Initial experiments were set up using the OpenShift
Platform-as-a-Service, which provides a free tier within which this
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. No special care was taken to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proved that the influence
of these islands in the next experiment is, in fact negligible.
%
\begin{table*}d
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline 
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|cccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt PUT}s & $<$ 69s & $<$ 3.46s \\
\hline 
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% \\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}. This table shows the median number of IPs
intervening in each experiment,  median of the time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment. The
maximum number of different IPs for each experiment is also in the
same ballpark and of the order of 10, which is also consistent with
prior work. The median time has a big range of variation, but 50\% of
the time takes less than several minutes, from around 4 minutes in the
best case to roughly 2/3 of an hour in the worst case. That figure is
not competitive, {\em a priori}, with the baseline experiment, that is
why it is interesting to look at the last two columns, that measure
the percentage of experiments in which the solution was found in less
time than the baseline. It is similar in all three cases, with around
20\% on average for the first baseline experiment, roughly 3\% in the
second case. The comparison is not totally fair: this experiment takes
place in a browser and in fact most of the time is devoted to
presenting the charts; the server is not the same since it is run in a
PaaS instead of a high-performance desktop computer, and it is also a
remote server, not a local server as in the baseline experiment. But
fair is fair, and we were looking for creating a framework that
allowed to run evolutionary computation experiments {\em faster} than
in our very own computer. And the results is that although it will
happen, one out of roughly 5 or roughly 20, depending on what baseline
experiment you choose, it will not happen always or even a majority of
the time.

>>>>

Initial experiments were set up using the OpenShift
Platform-as-a-Service, which provides a free tier within which this
experiment could be performed. That way, and as required, the hardware
and cloud cost for this experiment were zero. Experiments were
announced through a post in Twitter and other social networks, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the months of April and then by the beginning of August. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. No special care was taken to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proved that the influence
of these islands in the next experiment is, in fact negligible.
%
\begin{table*}d
\caption{Characteristics of the volunteer computing initial runs. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rr}
\hline
Date & \# experiments & Different IPs \\
\hline 
April 4th 4/4 & 57 & 191 \\
April 24th 4/24 &  231 & 559 \\
July 31th 7/31 & 97 & 179 \\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
\begin{table*}
\caption{Summary of time per run, number of IPS and number of {\tt
    PUT}s per IP in the initial runs. \label{tab:summary:os}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \#{\tt
  PUT}s & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline 
4/4 & 5 & 16 & 2040 & 18 & 14.29\% & 5.36\% & 0.0082 \\
4/24 &  5 & 29 & 732 & 11 & 47.39\% & 3.91\% & 0.0934\\
7/31 & 5 & 14 & 260 & 23 & 27.08\% & 1.04\%  & 0.1741\\
\hline 
\end{tabular}
\end{center}
\end{table*}
%
The table shows that every run included more than 50 experiments. The
number of different IPs intervening in them varied from more than one
hundred to more than five hundred in the second experiment.

A summary of the results of each run is also shown in Table
\ref{tab:summary:os}. This table shows the median number of IPs
intervening in each experiment,  median of the time needed
to finish the experiment, median number of HTTP {\tt PUT}s per IP, and
then two columns showing the percentage of experiments that took less
than the two baseline experiments shown in the introduction to this
section. The first striking result is that in all cases, 50\% of the
experiments involved 5 or less IPs. This is consistent with results
found previously \cite{DBLP:conf/gecco/GuervosG15} which found 6 to be
the usual number of IPs that participated in an experiment. The
maximum number of different IPs for each experiment is also in the
same ballpark and of the order of 10, which is also consistent with
prior work. The median time has a big range of variation, but 50\% of
the time takes less than several minutes, from around 4 minutes in the
best case to roughly 2/3 of an hour in the worst case. That figure is
not competitive, {\em a priori}, with the baseline experiment, that is
why it is interesting to look at the last two columns, that measure
the percentage of experiments in which the solution was found in less
time than the baseline. It is similar in all three cases, with around
20\% on average for the first baseline experiment, roughly 3\% in the
second case. The comparison is not totally fair: this experiment takes
place in a browser and in fact most of the time is devoted to
presenting the charts; the server is not the same since it is run in a
PaaS instead of a high-performance desktop computer, and it is also a
remote server, not a local server as in the baseline experiment. But
fair is fair, and we were looking for creating a framework that
allowed to run evolutionary computation experiments {\em faster} than
in our very own computer. And the results is that although it will
happen, one out of roughly 5 or roughly 20, depending on what baseline
experiment you choose, it will not happen always or even a majority of
the time. One of the problems is shown in the last column, which
indicates the statistical correlation between the number of IPs
participating in one experiment and the next. In all cases,
correlation is so low as to affirm that there is almost no
relationship between them, that is, computing nodes are not {\em
  staying} after the solution has been found; this is a feature of
this implementation: the user has to actively reload the page to make
the evolutionary algorithm start again. However, it puts a probably
unnecesary cap on the time, or number of experiments, that an user
will voluntarily perform.
